---
title: "ConnectorSharepoint"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ConnectorSharepoint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The connector.sharepoint package provides a convenient interface for accessing and interacting with Microsoft SharePoint sites directly from R.


## Introduction
This vignette will guide you through the process of connecting to a SharePoint site, retrieving data, and performing various operations using this package.

## Connecting to a SharePoint Site
### Without connector package

To get started, you need to establish a connection to your SharePoint site. Use
the `ConnectorSharepoint()` function to authenticate and connect to your SharePoint site. Here's an example of how to do this:

```{r, eval=FALSE}
library(connector.sharepoint)

# Connect to SharePoint
con <- connector_sharepoint(site_url = "sharepoint_url")
```

### With connector package (recommended)

If you are using the connector package, you can connect to a SharePoint site
using the `connect()` function. This function based on a configuration file or
a list creates a `connectors()` object with a `connector`for each of the
specified datasources (for detailed explanation have a look at connector package).

Configuration file for connecting to SharePoint should look like this:

```{r, eval=FALSE}

```

Save this to `_connector.yml` file and use the `connect()` function to connect
to SharePoint:

```{r, eval=FALSE}
library(connector)

# Create connector object
db <- connect()
```

Now you can access the SharePoint site using the `db` object and `adam` field
inside of it.

```{r, eval=FALSE}
# Connection to SharePoint site. This will print object details
db$adam
```

After the setup is done we can use this connection to manipulate Sharepoint data.

## Reading and writing data

The `connector` packages provide a set of functions to read and write data from
the datasources. They all have similar interface, so it's easy to switch between
them.

Now, we will show how to read and write different types data from/to Sharepoint.
In these examples we will be using `iris` and `mtcars` datasets.

Here is an example of writing data to the ADaM table:

```{r eval=FALSE}
library(dplyr)

# Manipulate data

## Iris data
setosa <- iris |>
  filter(Species == "setosa")

mean_for_all_iris <- iris |>
  group_by(Species) |>
  summarise_all(list(mean, median, sd, min, max))

## Mtcars data
cars <- mtcars |>
  filter(mpg > 22)

mean_for_all_mtcars <- mtcars |>
  group_by(gear) |>
  summarise(
    across(
      everything(),
      list("mean" = mean, "median" = median, "sd" = sd, "min" = min, "max" = max),
      .names = "{.col}_{.fn}"
    )
  ) |>
  tidyr::pivot_longer(
    cols = -gear,
    names_to = c(".value", "stat"),
    names_sep = "_"
  )

## Store data
db$adam |>
  write_cnt(setosa, "setosa", overwrite = TRUE)

db$adam |>
  write_cnt(mean_for_all_iris, "mean_iris", overwrite = TRUE)

db$adam |>
  write_cnt(cars, "cars_mpg", overwrite = TRUE)

db$adam |>
  write_cnt(mean_for_all_mtcars, "mean_mtcars", overwrite = TRUE)
```

Now, let's read the data back manipulate it a bit and write it to the SharePoint.
This way we can save different types of data in different formats.

```{r eval=FALSE}
library(gt)
library(tidyr)
library(ggplot2)

# List and load data
db$adam |>
  list_content_cnt()

table <- db$adam |>
  read_cnt("mean_mtcars")

gttable <- table |>
  gt(groupname_col = "gear")

# Save nontabular data to sharepoint
tmp_file <- tempfile(fileext = ".docx")
gtsave(gttable, tmp_file)
db$output |>
  upload_cnt(tmp_file, "tmeanallmtcars.docx")

# Manipulate data
setosa_fsetosa <- db$adam |>
  read_cnt("setosa") |>
  filter(Sepal.Length > 5)

fsetosa <- ggplot(setosa) +
  aes(x = Sepal.Length, y = Sepal.Width) +
  geom_point()

## Store data into output location
db$output |>
  write_cnt(fsetosa$data, "fsetosa.csv")
db$output |>
  write_cnt(fsetosa, "fsetosa.rds")

tmp_file <- tempfile(fileext = ".png")
ggsave(tmp_file, fsetosa)
db$output |>
  upload(tmp_file, "fsetosa.png")
```

## Conclusion

In this vignette we showed how to connect to SharePoint site, read and write
data from it. We also showed how to use the connector package to connect to
SharePoint and how to manipulate data using the connector package.
